---
output: github_document
---

```{r, echo = F}
library(simpopmods)

```

# R package: simpopmods

An R package for generating **sim**ulated **pop**ulation **mod**els (IPMs and MPMs)

:herb::mushroom::pig2::seedling::cactus::deciduous_tree::water_buffalo:

This package is under active development - do not use

## About

This is an R package I have been developing for my PhD research to simulate population models. It uses integral projection models.

## Installation

How to install this package from GitHub

```{r, eval = F}
install.packages("devtools")
library(devtools)
install_github("simonrolph/simpopmods")

```

## Usage

### Define an IPM

Before you generate any simulated populations, you first need to define a template for your population model. This is done through a IPM descriptor object. You can generate a template for the IPM descriptor by using the `init_IPM_desc()` function.

An `simpopmods` IPM descriptor consists of

 * Discrete states
 * Continous states (which all share the same domain)
 * Parameters
 * Functions for generating parameters that are not sampled directly
 * Demographic functions (equations of parameters)
 * Kernels (consisting of demographic functions)
 * Functions for the upper limit, lower limit and resolution of the kernel of the continious state domain.
 
Here's one I made earlier:
 
```{r}
load("~/PhD/simpopmods/IPM_descs/plant_basic_IPM_desc.Rdata")
```
 
### Create an IPM

Create a parameter set or use a sampler to sample parameter sets. I have been using an Adaptive Metropolis Algorithym with delayed rejection from the R package `FME`. This requires some informative priors.

```{r}
params <- c(
  surv.int  =  0,
  surv.z    =   2,
  flow.int  = 0.3,
  flow.z    =   0.1,
  grow.int  =   0.2,
  grow.sd   =   0.25,
  rcsz.sd   =   0.3,
  seed.int  =   2.35,
  seed.z    =   2.37,
  p.r       =   0.4
)

# create a complete parameter set
params_c <- make_full_params(params_ic = params, IPM_desc = IPM_desc)
```

Once you have a parameter set you can create an IPM and create the mega_kernels by:



```{r}
IPM <- make_kernels(params_c,IPM_desc)
```

This IPM can be discritised to an MPM by specifying a target stabe stage distribution for the MPM like so:

```{r}
qtiles <- c(0,0.2,0.4,0.7,1)
MPM <- make_MPM (params_c, IPM_desc, qtiles, submesh_res = 200)
```

lambda doesn't change in the discretisation process and you can use the `calc_dom_eig()` function to calculate the dominant eigenvalue from the mega kernel:

```{r, fig.height=7, fig.width=7, echo = F}
par(mfrow=c(2,2))
image(z=t(IPM$megakernels$P),x = IPM$meshpts,y = IPM$meshpts,main = "IPM growth/survival kernel",xlab = "size",ylab = "size at t+1")
image(z=t(MPM$megakernels$P),x = MPM$meshpts,y = MPM$meshpts,main = "MPM growth/survival kernel",xlab = "size",ylab = "size at t+1")
image(z=t(IPM$megakernels$Fec),x = IPM$meshpts,y = IPM$meshpts,main = "IPM fecundity kernel",xlab = "size",ylab = "size at t+1")
image(z=t(MPM$megakernels$Fec),x = MPM$meshpts,y = MPM$meshpts,main = "MPM fecundity kernel",xlab = "size",ylab = "size at t+1")
```

```{r, echo = F}
cat(paste("Lambda of IPM:",calc_dom_eig(IPM)$lambda))
cat(paste("Lambda of MPM:",calc_dom_eig(MPM)$lambda))
```

This R package has tests inplemented with R package `testthat`


## Related packages

 * [popbio](https://github.com/cstubben/popbio) - popbio is an R package for modeling population growth rates using age- or stage-classified matrix models. The package consists mainly of methods described in Hal Caswell's Matrix Population Models (2001) and Morris and Doak's Quantitative Conservation Biology (2002).
 * [Rage](https://github.com/jonesor/Rage)
 * [Rcompadre](https://github.com/jonesor/Rcompadre)

## IPM book

[Data-driven Modelling of Structured Populations](http://www.springer.com/gb/book/9783319288918) :star::star::star::star::star: 

